---
title: "minxue_A6"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, messages = FALSE,error=FALSE)
```

```{r}
library(tidyverse)
library(sf)
library(tigris)
library(leaflet)
library(censusapi)
library(readxl)
library(mapview)

Sys.setenv(CENSUS_KEY="60e044e0b171438ffa284a00c4a369df21caa10e")
```


```{r}
pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "MV",
    "AGEP",
    "HINCP",
    "TEN",
    "BLD",
    "YBL"
  )
)
saveRDS(pums_2019_1yr, "a6_pums.rds")
pums_2019_1yr <- readRDS("a6_pums.rds")

```


```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)


sf_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME =="San Francisco")

sf_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[sf_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

mapview(sf_pumas)
```


```{r}
sf_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% sf_pumas$PUMACE10)
```


```{r}
sf_pums_clean <- sf_pums %>% 
  mutate(
    YBL = as.numeric(YBL),
    AGEP = as.numeric(AGEP),
    HINCP = as.numeric(HINCP),
         )%>%
  filter(YBL %in% as.character(1:3)) %>% 
  arrange(AGEP) %>% 
  group_by(SERIALNO) %>% 
  summarize_all(first)
```

```{r}
sf_pums_lead <-
  sf_pums_clean %>% 
  mutate(
    lead_risk_population = ifelse(
      (AGEP< 6) & (HINCP< 90000),
      1,
      0
    ),
    income = as.numeric(HINCP)
  ) %>% 
  filter(
    income > 0
  )
```

```{r}
sf_pums_lead <- sf_pums_lead %>% 
 mutate(
    MV = MV %>% 
      factor(
        levels = sf_pums_lead$MV %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )
 )
sf_pums_lead <- sf_pums_lead %>% 
 mutate(
    TEN= TEN %>% 
      factor(
        levels = sf_pums_lead$TEN %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )
 )
sf_pums_lead <- sf_pums_lead %>% 
 mutate(
    BLD = BLD %>% 
      factor(
        levels = sf_pums_lead$BLD %>% 
          unique() %>%
          as.numeric() %>% 
          sort()
      )
 )
sf_pums_lead <- sf_pums_lead %>% 
 mutate(
    PUMA = PUMA %>% 
      factor(
        levels = sf_pums_lead$PUMA %>% 
          unique() %>%
          sort()
      )
 )
```


```{r}
logit_model <- glm(
  lead_risk_population ~ TEN + MV + BLD + PUMA,
  family = quasibinomial(),
  data = sf_pums_lead
)
summary(logit_model)
```

  The model treats the variable categories alphabetically, taking TEN1, MV1, BLD1, and PUMA07501 to be the “baselines” by which the other categories are compared. Controlling the explanatory power of other variables, the regression coefficients can be interpreted as differences in the specific category.
  In the Tenure category, compared with Ten1, which is "Owned with mortgage or loan (include home equity loans)", categories "4": "Occupied without payment of rent", and "3": "Rented", have higher lead risk while "2": "Owned Free And Clear" has lower lead risk. In the BLD category, compared with BLD1, Mobile Home or Trailer, all remaining types of homes are at higher lead risk, with "02": "One-family house detached" and "03": "One-family house attached" having the highest coefficients. In the MV category, compared with MV1: moved in 12 months or less, "3": moved in in 2 to 4 years" and "4": moved in in "5 to 9 years" have the highest coefficients. Lastly, in terms of spatial areas, households in census block PUMA07507 seem to have the highest lead risk.


```{r}
predict_dataframe=sample_n(sf_pums_lead, 1)
predict_dataframe
predict(logit_model, predict_dataframe, type = "response")
```



```{r}
lead_predicted <-
  predict(logit_model, data = sf_pums_lead, type = "response")
data=cbind(sf_pums_lead, lead_predicted)
```


```{r}
summary_2x2 <-
  data %>% 
  mutate(
    lead = ifelse(
      lead_risk_population == 1, 
      "Yes", 
      "No"
    )
  ) %>% 
  pull(lead_risk_population) %>% 
  table(lead_predicted > 0.1)
```

```{r}
summary_2x2
```

  The bottom-right cell is the number of households at risk for lead exposure with an income of less than 90,000 and at least a child under 6 years of age, and the model correctly predicted the result using variables of and BLD, TEN, MV, and PUMA as predictors. The top-left cell is the number of households we assumed not at risk of lead, and the model predicted the same result. So 97% of records were correctly predicted one way or the other. The top-right cell is the number of households we assumed not at risk of lead but the model incorrectly predicted them to be at risk which is “false positives”. The bottom-left cell is the number of households who are actually at risk for lead exposure, but the model incorrectly predicted them to be safes which is “false negatives”.
  According to the table, all 31 under the TRUE category should be mailed out based on the 0.1 threshold, of the total 31, 4 will go to right places which are low-income households with children, so the percentage is 4/31. Since the number of all at-risk households which display ‘1’ under the lead risk category mounts to 37, 4/37 of all at-risk children are likely to be reached. Possible ways to better address the problem is to lower the threshold for judging whether there is lead risk. Although it will introduce more “false positives” and cause a certain waste of resources, more families with real risk can be included which is the whole purpose of building the data set. In addition, according to the fitting relationship of Logic Model, more attention should be given to the categories with highest lead risk, such as providing more kits to household types of One-family houses.


