---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
```
```
```{r}
library(tidyverse)
year <- 2017:2020
quarters <- 1:4
type <- "Electric"
pge_total_elec <- NULL
for(quarter in quarters) {
  filename <- 
    paste0(
      "pge/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )
  print(filename)
  temp <- read_csv(filename)
  pge_total_elec <- rbind(pge_total_elec,temp)
  saveRDS(pge_total_elec, "pge_total_elec.rds")
}
  
year <- 2021
quarters <- 1:2
type <- "Electric"
pge_totalplus_elec <- pge_total_elec
for(quarter in quarters) {
  filename <- 
    paste0(
      "pge/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )
  print(filename)
  temp <- read_csv(filename)
  pge_totalplus_elec <- rbind(pge_totalplus_elec,temp)
  saveRDS(pge_totalplus_elec, "pge_totalplus_elec.rds")
}
years <- 2017:2020
quarters <- 1:4
type <- "Gas"
pge_total_gas <- NULL
for(year in years ){
  for(quarter in quarters) {
  filename <- 
    paste0(
      "pge/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )

  print(filename)
  temp <- read_csv(filename)
  pge_total_gas <- rbind(pge_total_gas,temp)
  saveRDS(pge_total_gas, "pge_total_gas.rds")
}
  }
year <- 2021
quarters <- 1:2
type <- "Gas"
pge_totalplus_gas <- pge_total_gas
for(quarter in quarters) {
  filename <- 
    paste0(
      "pge/PGE_",
      year,
      "_Q",
      quarter,
      "_",
      type,
      "UsageByZip.csv"
    )
  print(filename)
  temp <- read_csv(filename)
  pge_totalplus_gas <- rbind(pge_totalplus_gas,temp)
  saveRDS(pge_totalplus_gas, "pge_totalplus_gas.rds")
}
```

```{r}
pge_filterC_elec <- 
  filter(
    pge_totalplus_elec,
    CUSTOMERCLASS %in% 
      c(
        "Elec- Commercial"
      )
  )
pge_filterC_gas <- 
  filter(
    pge_totalplus_gas,
    CUSTOMERCLASS %in% 
      c(
        "Gas- Commercial"
      )
  )
pge_selectC_elec <-
  select(
    pge_filterC_elec,
    -c(COMBINED, AVERAGEKWH,TOTALCUSTOMERS)
  )
pge_selectC_gas <-
  select(
    pge_filterC_gas,
    -c(COMBINED, AVERAGETHM,TOTALCUSTOMERS)
  )
library(zoo)
library(tidyverse)
pge_mutateC_elec <-
  mutate(
    pge_selectC_elec,
    TOTALBTU =
      TOTALKWH * 3.412, 
    TIME=paste0(YEAR,'-',MONTH,"-1")
  )

pge_mutateC_gas <-
  mutate(
    pge_selectC_gas,
    TOTALBTU =
      TOTALTHM * 100, 
    TIME=paste0(YEAR,'-',MONTH,"-1")
  )
pge_groupC_elec <-
  group_by(
    pge_mutateC_elec,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TIME
  )
pge_groupC_gas <-
  group_by(
    pge_mutateC_gas,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TIME
  )
pge_summarizeC_elec <-
  summarize(
    pge_groupC_elec,
    TOTALBTU = 
      sum(
        TOTALBTU, 
        na.rm = T
      )
  )
pge_summarize_gas <-
  summarize(
    pge_groupC_gas,
    TOTALBTU = 
      sum(
        TOTALBTU, 
        na.rm = T
      )
  )
pge_finalC<-
rbind(pge_summarizeC_elec,pge_summarize_gas)

```

```{r}
pge_chart <-
  pge_finalC %>% 
  ggplot() +
  geom_bar(
    aes(
      x = as.Date(TIME, "%Y-%m-%d"),
      y = TOTALBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTU",
    title = "PG&E Territory Monthly Commercial Energy Usage, 2017-2021",
    fill = "Energy Type"
  )

pge_chart
```

```{r}
pge_filterR_elec <- 
  filter(
    pge_totalplus_elec,
    CUSTOMERCLASS %in% 
      c(
        "Elec- Residential"
      )
  )
pge_filterR_gas <- 
  filter(
    pge_totalplus_gas,
    CUSTOMERCLASS %in% 
      c(
        "Gas- Residential"
      )
  )
pge_selectR_elec <-
  select(
    pge_filterR_elec,
    -c(COMBINED, AVERAGEKWH,TOTALCUSTOMERS)
  )
pge_selectR_gas <-
  select(
    pge_filterR_gas,
    -c(COMBINED, AVERAGETHM,TOTALCUSTOMERS)
  )
pge_mutateR_elec <-
  mutate(
    pge_selectR_elec,
    TOTALBTU =
      TOTALKWH * 3.412,  TIME=paste0(YEAR,'-',MONTH,"-1"))

pge_mutateR_gas <-
  mutate(
    pge_selectR_gas,
    TOTALBTU =
      TOTALTHM * 100,  TIME=paste0(YEAR,'-',MONTH,"-1"))

pge_groupR_elec <-
  group_by(
    pge_mutateR_elec,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TIME
  )
pge_groupR_gas <-
  group_by(
    pge_mutateR_gas,
    MONTH,
    YEAR,
    CUSTOMERCLASS,
    TIME
  )
pge_summarizeR_elec <-
  summarize(
    pge_groupR_elec,
    TOTALBTU = 
      sum(
        TOTALBTU, 
        na.rm = T
      )
  )
pge_summarize_gas <-
  summarize(
    pge_groupR_gas,
    TOTALBTU = 
      sum(
        TOTALBTU, 
        na.rm = T
      )
  )
pge_finalR<-
rbind(pge_summarizeR_elec,pge_summarize_gas)

```

```{r}
pge_chart <-
  pge_finalR %>% 
  ggplot() +
  geom_bar(
    aes(
      x = as.Date(TIME, "%Y-%m-%d"),
      y = TOTALBTU,
      fill = CUSTOMERCLASS
    ),
    stat = "identity",
    position = "stack"
  ) +
  labs(
    x = "Month",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Energy Usage, 2017-2021",
    fill = "Energy Type")
pge_chart
```
```{r}
pge_20192020C<- 
  filter(pge_finalC, YEAR%in% 
      c(
        "2019",
        "2020"
      )  )
pge_20192020CElec=filter(pge_20192020C, CUSTOMERCLASS %in% 
      c(
        "Elec- Commercial"
      ) )
pge_20192020CGas=filter(pge_20192020C, CUSTOMERCLASS %in% 
      c(
        "Gas- Commercial"
      ) )
```


```{r}
library(ggplot2)
p20192020C<- ggplot() + geom_line(data=pge_20192020CGas,aes(x = as.Date(TIME, "%Y-%m-%d") ,y = TOTALBTU,color="red"))+ geom_line(data=pge_20192020CElec,aes(x = as.Date(TIME, "%Y-%m-%d") ,y = TOTALBTU, color="blue")) +labs(
    x = "Month",
    y = "kBTU",
    title = "PG&E Territory Monthly Commercial Energy Usage, 2019-2020",
  )+scale_color_discrete(name = "Energy use", labels = c("20192020REle", "20192020RGas"))
p20192020C

```

Caveats about the data sources：We need to determine whether and to what extent differences in energy use are due to weather. If there is a significant difference in temperature between the two years, such as extreme heat in summer and low temperatures in winter, it could offset or aggravate the performance of the corona virus on electricity demand.

There are strong positive correlations between the progression of the pandemic and the commercial electricity consumption.The commercial sector experienced a visible decrease in electricity usage, showing pronounced differences in the electricity usage the year before COVID and at the start of COVID, which is March 2020.Many businesses were operating at partial capacity and experienced a sharp decline in customers for shopping in stores, thus reducing their demand for electricity, including cooling lighting and sound equipment.
COVID-19 pandemic has had less effect on the demand for gas than for electricity.Consumption of natural gas fell，especially in the first quarter，as consumer services such as restaurants, culture and entertainment and transportation temporarily suspended operations.


```{r}
pge_20192020R<- 
  filter(pge_finalR, YEAR%in% 
      c(
        "2019",
        "2020"
      )  )
pge_20192020RElec=filter(pge_20192020R, CUSTOMERCLASS %in% 
      c(
        "Elec- Residential"
      ) )
pge_20192020RGas=filter(pge_20192020R, CUSTOMERCLASS %in% 
      c(
        "Gas- Residential"
      ) )
```


```{r}
library(ggplot2)
p20192020R<- ggplot() + geom_line(data=pge_20192020RGas,aes(x = as.Date(TIME, "%Y-%m-%d") ,y = TOTALBTU,color="red"))+ geom_line(data=pge_20192020RElec,aes(x = as.Date(TIME, "%Y-%m-%d"),y = TOTALBTU,color="blue")) +labs(
    x = "Month",
    y = "kBTU",
    title = "PG&E Territory Monthly Residential Energy Usage, 2019-2020",
  )+scale_color_discrete(name = "Energy use", labels = c("20192020REle", "20192020RGas"))
p20192020R
```

The epidemic led to a short-term sharp increase in electricity consumption in the residential sector in March in 2020, and the increase gradually slowed down.This is due to at the beginning of the COVID-19 pandemic, people were isolated at home, and employees of the company had to work online. Household electricity demand for heating, lighting, TV, computer and other aspects increased, and household electricity consumption kept increasing. 
In addition to the continuous impact of COVID-19 isolation on residential electricity consumption, temperature has a greater impact on residential electricity consumption from March to August, with July being the peak season, which may be due to the need to use electricity for cooling such as air conditioning in summer. To sum up, residential electricity consumption in 2020 increased compared to 2019.
Gas use is roughly the same as before COVID-19, with a slight increase in march when quarantine rules set in.Subsequent consistency with the previous year may be due to people getting used to the impact of COVID-19。

