---
title: "Assignment 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, messages = FALSE,error=FALSE)
```     
```{r}
library(tigris)
library(tidyverse)
library(tidycensus)
library(sf)
library(censusapi)
library(leaflet)
library(StatMatch)
library(mapview)
library(readxl)
library(spdep)
library(spatialreg)
library(survey)
Sys.setenv(CENSUS_KEY="60e044e0b171438ffa284a00c4a369df21caa10e")
```

```{r}
ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()
```
```{r}
leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = bay_pumas,
    weight = 1,
    color = "gray",
    label = ~PUMACE10
  ) %>% 
  addMarkers(
    lng = -121.78059339592423,
    lat = 37.997870545675404
  ) %>% 
  addPolygons(
    data = bay_pumas %>% 
      filter(PUMACE10 == "01309 ")
  )
```

```{r}
pums_2014_2019 <- readRDS("pums_2014_2019_wts.rds")
```

```{r}
pums_bart <- pums_2014_2019 %>%
  mutate(
    PWGTP = as.numeric(PWGTP),
    bart = ifelse(
      JWTR %in% c("4"),
      PWGTP,
      0
    )
  ) %>% 
  group_by(PUMA, year) %>% 
  summarize(
    pop = sum(PWGTP),
    bart = sum(bart)
  )
```



```{r}
pums_bart_clean <-
  pums_bart %>% 
  select(-pop) %>% 
  pivot_wider(
    names_from = year,
    values_from = bart
  )
```

```{r}
obs_matrix <-
  pums_bart_clean %>% 
  ungroup() %>% 
  select(`2014`,`2015`,`2016`,`2017`) %>% 
  as.matrix()

dist_matrix <- mahalanobis.dist(obs_matrix)

rownames(dist_matrix) <- pums_bart_clean$PUMA
colnames(dist_matrix) <- pums_bart_clean$PUMA

match <- dist_matrix["01309",] %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(
    PUMA = rowname,
    match = "."
  ) %>% 
  right_join(
    pums_bart_clean
  ) %>% 
  arrange(match) %>% 
  .[1:11, ] %>% 
  left_join(bay_pumas %>% select(PUMA = PUMACE10)) %>% 
  st_as_sf()
```

```{r}
leaflet() %>% 
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addPolygons(
    data = match[1, ],
    color = "red",
    label = ~PUMA
  ) %>% 
  addPolygons(
    data = match[-1, ],
    label = ~PUMA
  )
```

```{r}
match_pumas <-
  match %>% 
  filter(!PUMA %in% c("01309")) %>% 
  st_drop_geometry() %>% 
  select(-match) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  ) %>%
  group_by(
    year
  ) %>% 
  summarize(
    bart = mean(bart),
    PUMA = "Similar PUMAs"
  )

treatment_pumas <-
  match %>% 
  filter(PUMA %in% c("01309")) %>% 
  select(-match) %>% 
  st_drop_geometry() %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  )

rbind(
  treatment_pumas,
  match_pumas
) %>% 
  ggplot(
    aes(
      x = as.numeric(year),
      y = bart,
      color = PUMA
    )
  ) +
  geom_line() +
  geom_vline(xintercept = 2018, linetype = "dashed") +
  labs(
    title = "Milpitas vs. control neighborhoods, BART ridership",
    x = "Year",
    y = "BART commute riders"
  )
```
```{r}
transit_did <-
  match %>% 
  st_drop_geometry() %>% 
  select(-match) %>% 
  pivot_longer(
    -PUMA,
    names_to = "year",
    values_to = "bart"
  ) %>% 
  mutate(
    year = year %>% as.numeric(),
    time = ifelse(year >= 2018, 1, 0),
    treated = ifelse(PUMA == "01309", 1, 0)
  )

did_reg <- lm(bart ~ treated*time, data = transit_did)

summary(did_reg)
```


Assumptions:
We chose 10 matching PUMAs using Mahalanobis distance.
The BART station opened in mid-2018, and the PUMS used in the analysis included four years before and two years after treatment, in part because of the limited amount of data.
We created the logit model comparing one Puma as opposed to the other ones by encoding that time is only treated if time is after 2018, and the area was only treated if it is 01309.
The PUMA data used for the statistics draws from pool of “Subway or elevated car” in the ACS and interpetated BART as a mode of work travel.

Result:
The red line and green line before 2018 visualize the parallel trends assumption we made. Green line is the baseline to rule out the possibility that there is a general increase in subway transportation. There is a spike between 2017 and 2018 of the red line and the difference between the red and the green line after 2018 appears to be the unique contribution of the new Bart station.

In the model:
Treated represents the pre-treatment difference between treatment and control PUMAs and indicate the difference between rea and green lines before the treatment (viewable in the previous line graph as the average vertical difference between the two lines before 2018). According to the summary, this vertical differences is 471.
Time represents the change in the control PUMAs from pre-treatment to post-treatment, which is the difference of the average of the Green Line in the first four and the next two years. According to the summary , this differences is 207.28.
Treated asterisk time creates interaction effect between the two, and the result appears to be statistically significant (p-value less than 5%), stating the Milpitas BART station in 2018, as having had an estimated impact of about 1100 new BART commuters. 



```{r}
pums_2014_2019 <- readRDS("pums_2014_2019_wts.rds")
pums_2014_2019 <- pums_2014_2019 %>%   
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")) %>%   
right_join(transit_did)%>%
  group_by(
    PUMA,year
  ) %>% summarise_all(first)%>%filter_all(all_vars(!is.na(.))) 

```

```{r}
design <- svrepdesign(
  data = pums_2014_2019,
  type = "ACS",
  repweights = pums_2014_2019[ ,9:88],
  weights = ~as.numeric(PWGTP)
)
```


```{r}
model <- svyglm(
  formula = bart ~ treated*time,
  family = gaussian(),
  design = design
)

summary(model)
```

After doing the replicate weights, the Estimate remains roughly the same while Std. Error t value is significantly smaller.
